---
title: "Reproducible Research: Peer Assessment 1"
author: Jurij Robba
output: 
html_document:
keep_md: true
---


## Loading and preprocessing the data

Data is read from file, steps are transformed to numeric:

```{r data, include=TRUE}

#readout
data <- read.csv("activity.csv")

#to numeric
data$steps <- as.numeric(data$steps)

```

## What is mean total number of steps taken per day?

First we aggregate data and create histogram:

```{r meanhist, include=TRUE}


#agregating
daily_sum <- aggregate(data$steps, by=list(data$date), FUN='sum', na.rm=TRUE)

#histogram
hist(daily_sum$x, xlab="number of steps", main="Frequency of number of steps ber day")

```

We procede by taking mean:

```{r mean, include=TRUE}

#mean
mean(daily_sum$x)

```

and than median:

```{r median, include=TRUE}

#median
median(daily_sum$x)

```

## What is the average daily activity pattern?

We agregate and plot data:

```{r daily, include=TRUE}

#aggregate
interval_mean <- aggregate(data$steps, by=list(data$interval), FUN='mean', na.rm=TRUE)

#plot
plot(interval_mean$Group.1, interval_mean$x, xlab='interval (min)', ylab='steps', main='Average number of steps in interval', type='l')

```

Finding interval with most steps taken on average:

```{r max, include=TRUE}

interval_mean[which(interval_mean$x==max(interval_mean$x)),1]

```

This is at 13:55 - 14:00.

## Imputing missing values

To count rows with Na in original dataset we use:

```{r NaRows, include=TRUE}

nrow(na.omit(data))

```

Wherever data is missing we'll fill it with mean for interval:

```{r Fillmissing, include=TRUE}

interpolated_data <- data
missing_intervals <- interpolated_data[is.na(interpolated_data$steps),3]
missing_steps <- sapply(missing_intervals, function(x) { interval_mean[interval_mean$Group.1==x,2] })
interpolated_data[is.na(interpolated_data$steps),1] <- missing_steps

```

We calculate new histogram:

```{r newhist, include=TRUE}

#agregating
daily_sum <- aggregate(interpolated_data$steps, by=list(interpolated_data$date), FUN='sum', na.rm=TRUE)

#histogram
hist(daily_sum$x, xlab="number of steps", main="Frequency of number of steps ber day")

```

We procede by taking new mean:

```{r newmean, include=TRUE}

#mean
mean(daily_sum$x)

```

and than new median:

```{r newmedian, include=TRUE}

#median
median(daily_sum$x)

```

Both median and mean are higher now.

## Are there differences in activity patterns between weekdays and weekends?

Creating new variable in data

```{r weekend, include=TRUE}

data$weekend[weekdays(as.Date(data$date)) == "Sunday" | weekdays(as.Date(data$date)) == "Saturday"] <- 'weekend'
data$weekend[!(weekdays(as.Date(data$date)) == "Sunday" | weekdays(as.Date(data$date)) == "Saturday")] <- 'weekday'
data$weekend <- as.factor(data$weekend)

```

Creating panel plot

```{r panelplot, include=TRUE}

interval_mean_weekend <- aggregate(data$steps[data$weekend=='weekend'], by=list(data$interval[data$weekend=='weekend']), FUN='mean', na.rm=TRUE)
interval_mean_weekday <- aggregate(data$steps[data$weekend=='weekday'], by=list(data$interval[data$weekend=='weekday']), FUN='mean', na.rm=TRUE)

par(mfrow=c(2, 1), mar=c(4,4,2,1))

plot(interval_mean_weekend, type='l', xlab='', ylab='steps', main='weekend')
plot(interval_mean_weekday, type='l', xlab='interval', ylab='steps', main='weekdays')

```


